---
title: "Project 4"
output:
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: fill
    runtime: shiny
    theme:
      base_font:
        google: Prompt
      primary: '#34a4eb'
resource_files:
- images/coffee_beans.png
---

```{r setup, include=FALSE}
library(flexdashboard)
library(tidytuesdayR)
library(ggplot2)
library(tidyr)
library(dplyr)
library(ggalt)
library(purrr)
library(scales)
library(DT)

if(!dir.exists("data")){
  dir.create("data")
}

if(!file.exists("data/coffee.RDS")) {
  tuesdata <- tt_load("2020-07-07")
  coffee <- tuesdata$coffee_ratings

  saveRDS(coffee, file = "data/coffee.RDS")
}

coffee <- readRDS("data/coffee.RDS")
```
# About
The question I'm hoping to answer is: What factors are associated with high graded coffee beans?

The data comes from TidyTuesday ([2020-07-07](https://github.com/rfordatascience/tidytuesday/blob/master/data/2020/2020-07-07/readme.md)) The data dictionary is located [here](https://github.com/rfordatascience/tidytuesday/blob/master/data/2020/2020-07-07/readme.md#coffee_ratingscsv).

###
![[Source: TidyTuesday](https://github.com/rfordatascience/tidytuesday/blob/master/data/2020/2020-07-07/readme.md)](images/coffee_beans.png)

# The Data
The data originally comes from the Coffee Quality Institute. The data is located [here](https://github.com/jldbc/coffee-quality-database). Ratings and other scores come from trained reviewers from Coffee Quality Institute. The overall coffee score is the sum of all qualities measures (aroma, flavor, afteraste, etc.).

###
```{r}
coffee_table <- coffee %>%
  select(-c("altitude_mean_meters", "altitude_high_meters", "altitude_low_meters", "unit_of_measurement", "certification_contact", "certification_address", "lot_number", "ico_number", "in_country_partner", "owner_1", "variety", "expiration")) %>%
  rename("Coffee Rating" = "total_cup_points",
         "Species" = "species",
         "Owner" = "owner",
         "Country of Origin" = "country_of_origin",
         "Farm Name" = farm_name,
         "Mill" = "mill",
         "Company" = "company",
         "Altitude" = "altitude",
         "Region" = "region",
         "Producer" = "producer",
         "Number of Bags" = "number_of_bags",
         "Bag Weight" = "bag_weight",
         "Harvest Year" = "harvest_year",
         "Grading Date" = "grading_date",
         "Processing Method" = "processing_method",
         "Aroma" = "aroma",
         "Flavor" = "flavor",
         "Aftertaste" = "aftertaste",
         "Acidity" = "acidity",
         "Body" = "body",
         "Balance" = "balance",
         "Uniformity" = "uniformity",
         "Clean Cup" = "clean_cup",
         "Sweetness" = "sweetness",
         "Cupper Points" = "cupper_points",
         "Moisture" = "moisture",
         "Category 1 Defects" = "category_one_defects",
         "Quakers" = "quakers",
         "Color" = "color",
         "Category 2 Defects" = "category_two_defects",
         "Certification Body" = "certification_body")
datatable(coffee_table)
```

# Coffee Rating Distribution (Static)

```{r}
ggplot(coffee, aes(x = total_cup_points)) +
  geom_histogram() +
  labs(title = "Distribution of Coffee Ratings",
       x = "Coffee Ratings",
       y = "Rating Distribution")
```

# Coffee Bean Color Rating (Static)
```{r}
bean_color <- coffee %>%
  select(total_cup_points, color) %>%
  drop_na()

ggplot(bean_color, aes(x = color, y = total_cup_points)) +
  geom_boxplot() +
  labs(title = "Coffee Rating by Bean Color",
       x = "Coffee Bean Color",
       y = "Coffee Rating Distribution")
```

# Country of Origin Rating (Interactive)
## Sidebar {.sidebar}
```{r}
country_names <- unique(coffee$country_of_origin)
selectInput("c_name", label = "Country of Origin:",
            choices = country_names, selected = "Brazil")
```

## Chart
```{r}
renderPlot ({
  ggplot(coffee %>%
           distinct() %>%
           select(total_cup_points, country_of_origin) %>%
           filter(country_of_origin == input$c_name) %>%
           drop_na(), 
         aes(x = total_cup_points)) +
    geom_density() +
    labs(x = "Coffee Ratings",
         y = "Density",
         title = "Coffee Rating Denisty",
         subtitle = "By Country of Origin")
})

```

# Coffee Rating Factors (Interactive)
## Sidebar {.sidebar}
```{r}
selectInput("rate_fct", label = "Rating Factor:",
            choices = c("aroma", "flavor", "aftertaste", "acidity", "body", "balance"), selected = "aroma")
```

## Chart
```{r}
coffct_subtitle <- reactive({
    paste("Coffee Factor: ", input$rate_fct, sep = "")
  })

coff_fct_df <- coffee %>%
  select(total_cup_points, aroma, flavor, aftertaste, acidity, body, balance)

renderPlot ({
  ggplot(coff_fct_df, aes(x = !!as.symbol(input$rate_fct), y = total_cup_points)) +
    geom_point() +
    labs(x = "Factor Score",
         y = "Coffee Rating",
         title = "Coffee Ratings by Coffee Factor Score",
         subtitle = coffct_subtitle())
})

```

# Analysis
## Column {data-width="600"}

### Flavor Grade and Coffee Rating
```{r}
coffee_flv_bal <- coffee %>%
  distinct() %>%
  select(total_cup_points, flavor, balance, color) %>%
  filter(balance > 5.0) %>%
  drop_na() %>%
  group_by(flavor, color) %>%
  summarise(rating_means = mean(total_cup_points))

ggplot(coffee_flv_bal, aes(x = flavor, y = rating_means)) +
  facet_wrap(vars(color)) +
  geom_point() +
  geom_smooth(se=FALSE, linetype="dashed", size=0.5) +
  geom_xspline(size=0.5) +
  labs(x = "Flavor Grade",
       y = "Overall Rating",
       title = "Coffee Ratings by Flavor Ratings",
       subtitle = "By Color of Bean")
```

## Column {data-width="400"}

### Coffee Rating and Certification Body

```{r}
coffee_cert_rate <- coffee %>%
  select(total_cup_points, certification_body) %>%
  group_by(certification_body) %>%
  summarise(rating_means = mean(total_cup_points))

coffee_cert_rate$percent <- as.double(map(coffee_cert_rate$rating_means, function(x) x/100))



ggplot(coffee_cert_rate, aes(y = reorder(certification_body, percent), x = percent)) +
  geom_lollipop(horizontal = TRUE) +
  theme(axis.text.y = element_text(size=5, angle=30)) +
  scale_y_discrete(label = abbreviate) +
  labs(x = "Percent Coffee Rating",
       y = "Cerficiation Body",
       title = "Coffee Ratings by Certification Bodies",
       subtitle = "A Lollipop Chart")
```

### Coffee Rating and Defects

```{r}
defects <- coffee %>%
  select(category_one_defects, category_two_defects) %>%
  drop_na()

defect_total <- pmap_dbl(defects, .f = sum)

defects$dfct_tot <- defect_total

coffee_dfct <- data.frame(defects, coffee$total_cup_points)

ggplot(coffee_dfct, aes(x = dfct_tot, y = coffee.total_cup_points)) + 
  geom_point() +
  geom_smooth(se=FALSE, linetype="dashed", size=0.5) +
  labs(x = "Total Defects",
       y = "Overall Rating",
       title = "Coffee Rating by Total Defects",
       subtitle = "Including Primary and Secondary Defects")
```
