---
title: "Project 4"
output:
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: fill
    runtime: shiny
    theme:
      base_font:
        google: Prompt
      primary: '#237a94'
resource_files:
- images/coffee_beans.png
---

```{r setup, include=FALSE}
library(flexdashboard)
library(tidytuesdayR)
library(ggplot2)
library(tidyr)
library(dplyr)
library(ggalt)
library(purrr)
library(scales)
library(DT)

if(!dir.exists("data")){
  dir.create("data")
}

if(!file.exists("data/coffee.RDS")) {
  tuesdata <- tt_load("2020-07-07")
  coffee <- tuesdata$coffee_ratings

  saveRDS(coffee, file = "data/coffee.RDS")
}

coffee <- readRDS("data/coffee.RDS")
```
# About
This data looks at the ratings of various coffee beans. It includes various factors like flavor and aroma. I thought it would be interesting to see what factors go into high quality coffee.

The question I'm hoping to answer is: What factors are associated with high graded coffee beans?

The data comes from TidyTuesday ([2020-07-07](https://github.com/rfordatascience/tidytuesday/blob/master/data/2020/2020-07-07/readme.md)) The data dictionary is located [here](https://github.com/rfordatascience/tidytuesday/blob/master/data/2020/2020-07-07/readme.md#coffee_ratingscsv).

###
![[Source: TidyTuesday](https://github.com/rfordatascience/tidytuesday/blob/master/data/2020/2020-07-07/readme.md)](images/coffee_beans.png)

# The Data
The data originally comes from the Coffee Quality Institute. The data is located [here](https://github.com/jldbc/coffee-quality-database). Ratings and other scores come from trained reviewers from Coffee Quality Institute. The overall coffee score is the sum of all qualities measures (aroma, flavor, afteraste, etc.).

###
```{r}
coffee_table <- coffee %>%
  select(-c("altitude_mean_meters", "altitude_high_meters", "altitude_low_meters", "unit_of_measurement", "certification_contact", "certification_address", "lot_number", "ico_number", "in_country_partner", "owner_1", "variety", "expiration")) %>%
  rename("Coffee Rating" = "total_cup_points",
         "Species" = "species",
         "Owner" = "owner",
         "Country of Origin" = "country_of_origin",
         "Farm Name" = farm_name,
         "Mill" = "mill",
         "Company" = "company",
         "Altitude" = "altitude",
         "Region" = "region",
         "Producer" = "producer",
         "Number of Bags" = "number_of_bags",
         "Bag Weight" = "bag_weight",
         "Harvest Year" = "harvest_year",
         "Grading Date" = "grading_date",
         "Processing Method" = "processing_method",
         "Aroma" = "aroma",
         "Flavor" = "flavor",
         "Aftertaste" = "aftertaste",
         "Acidity" = "acidity",
         "Body" = "body",
         "Balance" = "balance",
         "Uniformity" = "uniformity",
         "Clean Cup" = "clean_cup",
         "Sweetness" = "sweetness",
         "Cupper Points" = "cupper_points",
         "Moisture" = "moisture",
         "Category 1 Defects" = "category_one_defects",
         "Quakers" = "quakers",
         "Color" = "color",
         "Category 2 Defects" = "category_two_defects",
         "Certification Body" = "certification_body")
datatable(coffee_table)
```

# Rating Distribution (Static)

```{r}
ggplot(coffee, aes(x = total_cup_points)) +
  geom_histogram() +
  labs(title = "Distribution of Coffee Ratings",
       x = "Coffee Ratings",
       y = "Rating Distribution")
```

# Coffee Bean Color (Static)
```{r}
bean_color <- coffee %>%
  select(total_cup_points, color) %>%
  drop_na()

ggplot(bean_color, aes(x = color, y = total_cup_points)) +
  geom_boxplot() +
  labs(title = "Coffee Rating by Bean Color",
       x = "Coffee Bean Color",
       y = "Coffee Rating Distribution")
```

# Country of Origin (Interactive)
## Sidebar {.sidebar}
```{r}
country_names <- unique(coffee$country_of_origin)
selectInput("c_name", label = "Country of Origin:",
            choices = country_names, selected = "Brazil")
```

## Chart
```{r}
renderPlot ({
  ggplot(coffee %>%
           distinct() %>%
           select(total_cup_points, country_of_origin) %>%
           filter(country_of_origin == input$c_name) %>%
           drop_na(), 
         aes(x = total_cup_points)) +
    geom_density() +
    labs(x = "Coffee Ratings",
         y = "Density",
         title = "Coffee Rating Denisty",
         subtitle = "By Country of Origin")
})

```

# Coffee Factors (Interactive)
## Sidebar {.sidebar}
```{r}
selectInput("rate_fct", label = "Rating Factor:",
            choices = c("aroma", "flavor", "aftertaste", "acidity", "body", "balance"), selected = "aroma")
```

## Chart
```{r}
coffct_subtitle <- reactive({
    paste("Coffee Factor: ", input$rate_fct, sep = "")
  })

coff_fct_df <- coffee %>%
  select(total_cup_points, aroma, flavor, aftertaste, acidity, body, balance)

renderPlot ({
  ggplot(coff_fct_df, aes(x = !!as.symbol(input$rate_fct), y = total_cup_points)) +
    geom_point() +
    labs(x = "Factor Score",
         y = "Coffee Rating",
         title = "Coffee Ratings by Coffee Factor Score",
         subtitle = coffct_subtitle())
})

```

# Analysis
## Column {data-width="600"}

### Chart 1: Flavor Grade and Coffee Rating
```{r}
coffee_flv_bal <- coffee %>%
  distinct() %>%
  select(total_cup_points, flavor, balance, color) %>%
  filter(balance > 5.0) %>%
  drop_na() %>%
  group_by(flavor, color) %>%
  summarise(rating_means = mean(total_cup_points))

ggplot(coffee_flv_bal, aes(x = flavor, y = rating_means)) +
  facet_wrap(vars(color)) +
  geom_point() +
  geom_smooth(se=FALSE, linetype="dashed", size=0.5) +
  geom_xspline(size=0.5) +
  labs(x = "Flavor Grade",
       y = "Overall Rating",
       title = "Coffee Ratings by Flavor Ratings",
       subtitle = "By Color of Bean")
```

## Column {data-width="400"}

### Chart 2: Coffee Rating and Certification Body

```{r}
coffee_cert_rate <- coffee %>%
  select(total_cup_points, certification_body) %>%
  group_by(certification_body) %>%
  summarise(rating_means = mean(total_cup_points))

coffee_cert_rate$percent <- as.double(map(coffee_cert_rate$rating_means, function(x) x/100))



ggplot(coffee_cert_rate, aes(y = reorder(certification_body, percent), x = percent)) +
  geom_lollipop(horizontal = TRUE) +
  theme(axis.text.y = element_text(size=5, angle=30)) +
  scale_y_discrete(label = abbreviate) +
  labs(x = "Percent Coffee Rating",
       y = "Cerficiation Body",
       title = "Coffee Ratings by Certification Bodies",
       subtitle = "A Lollipop Chart")
```


### Chart 3: Coffee Rating and Defects

```{r}
defects <- coffee %>%
  select(category_one_defects, category_two_defects) %>%
  drop_na()

defect_total <- pmap_dbl(defects, .f = sum)

defects$dfct_tot <- defect_total

coffee_dfct <- data.frame(defects, coffee$total_cup_points)

ggplot(coffee_dfct, aes(x = dfct_tot, y = coffee.total_cup_points)) + 
  geom_point() +
  geom_smooth(se=FALSE, linetype="dashed", size=0.5) +
  labs(x = "Total Defects",
       y = "Overall Rating",
       title = "Coffee Rating by Total Defects",
       subtitle = "Including Primary and Secondary Defects")
```


# Summary of Analysis
Looking at Chart 1, while the blue-ish green and uncolored beans have a generally linear increase in bean rating as flavor grade increases, the blue-green beans seem to increase drastically in rating at around a flavor grade of 6.5-7 (out of 10), while the green beans seem to have a dip in overall rating around a flavor grade of 6 and 7.

Looking at Chart 2, generally, it doesn't seem that the certification body affects a coffee bean's rating too much. METAD Agricultural Development plc seems to rate coffee beans the highest on average, and Instuto Hondereno del Cafe seems to rate coffee bean's the lowest on average (abbreviations used for the sake of space).

According to Chart 3, it seems the generally, the more defects a bean has the lower its rating, which is to be expected.